---
title: 'Mirai backend: fast, scalable parallel execution'
output: rmarkdown::html_vignette
vignette: '%\VignetteIndexEntry{7. Parallel Execution: mirai backend for scalable
  workflows} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}'
params:
  family: red
css: albers.css
resource_files:
- albers.css
- albers.js
includes:
  in_header: |-
    <script src="albers.js"></script>
    <script>document.addEventListener('DOMContentLoaded',()=>document.body.classList.add('palette-red'));</script>

---

```{r setup, include = FALSE}
if (requireNamespace("ggplot2", quietly = TRUE) && requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
if (requireNamespace("ggplot2", quietly = TRUE) && requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Why mirai?

The mirai backend brings several powerful advantages to parade workflows:

- **No connection limits**: Scale beyond R's 125 connection limit
- **Low-latency execution**: Persistent daemons with dispatcher for minimal overhead
- **Secure remote execution**: Built-in TLS and SSH tunneling support
- **HPC compliance**: Launch daemons through SLURM while maintaining scheduler policies
- **Simple local development**: Drop-in replacement for multisession with better performance

## Quick Start

### Local development

The simplest way to use mirai is as a replacement for the standard multisession backend:

```{r}
library(parade)

# Initialize mirai with auto-detected cores
mirai_init()

# Or use the distribution interface
fl <- flow(grid) |>
  stage("process", function(x) x^2, schema = returns(y = dbl())) |>
  distribute(use_mirai_local()) |>
  collect()
```

This provides instant parallelism without any configuration, and often with better performance than traditional backends for many small tasks.

### Comparison with other backends

| Scenario | Traditional | Mirai | Benefits |
|----------|-------------|--------|----------|
| Many small tasks | `dist_local()` | `use_mirai_local()` | Lower overhead, faster |
| >125 workers | Not possible | `dist_mirai(n = 200)` | No connection limits |
| Remote execution | Complex setup | `use_mirai_ssh()` | Built-in SSH tunneling |
| HPC clusters | `dist_slurm()` | `use_mirai_slurm()` | Lower latency, persistent workers |

## Distribution Patterns

### 1. Local Daemons

For development and single-machine workflows:

```{r}
# Basic local setup
dist <- dist_mirai(n = 8)

# With explicit configuration
dist <- dist_mirai(
  n = parallel::detectCores(),
  dispatcher = TRUE,  # Enable load balancing
  within = "mirai",   # Use mirai for nested parallelization
  workers_within = 4  # Nested workers per daemon
)

# Apply to flow
fl |> distribute(dist) |> collect()
```

### 2. SLURM-Managed Daemon Pools

Launch mirai daemons through SLURM for HPC-compliant distributed execution:

```{r}
# Configure SLURM-launched daemons
dist <- use_mirai_slurm(
  n = 16,                    # Number of daemon jobs
  partition = "compute",     # SLURM partition
  time = "2:00:00",         # Wall time
  mem = "32G",              # Memory per daemon
  cpus = 8,                 # CPUs per daemon
  tls = TRUE,               # Use TLS encryption
  port = 5555               # TLS port
)

# Submit flow with SLURM daemons
fl |> 
  distribute(dist) |>
  submit()  # Returns immediately

# Monitor status
deferred_status(handle)
results <- deferred_collect(handle)
```

The daemons run as SLURM jobs, ensuring compliance with cluster policies while providing low-latency task execution.

### 3. SSH Tunneling for Firewalled Clusters

Connect to remote nodes through SSH, even when direct connections are blocked:

```{r}
# SSH with tunneling (for firewalled nodes)
dist <- use_mirai_ssh(
  remotes = c("ssh://node1", "ssh://node2", "ssh://node3"),
  tunnel = TRUE,      # Use SSH tunneling
  port = 40491       # Local port for tunnel
)

# Direct SSH (when ports are open)
dist <- use_mirai_ssh(
  remotes = c("ssh://compute1", "ssh://compute2"),
  tunnel = FALSE
)

fl |> distribute(dist) |> collect()
```

SSH tunneling allows secure connections even in restricted network environments.

## Advanced Configuration

### Manual Daemon Setup

For full control over the mirai configuration:

```{r}
library(mirai)
library(future)

# Set up custom TLS-secured daemons
url <- host_url(tls = TRUE, port = 5555)

# Configure SLURM submission
opts <- paste(
  "#SBATCH --partition=highmem",
  "#SBATCH --time=04:00:00",
  "#SBATCH --mem=64G",
  "#SBATCH --job-name=my-analysis",
  sep = "\n"
)

cfg <- cluster_config(command = "sbatch", options = opts)

# Launch daemons
daemons(n = 32, url = url, remote = cfg, dispatcher = TRUE)

# Use with parade via future
plan(future.mirai::mirai_cluster)
fl |> collect()
```

### Dynamic Scaling

Adjust daemon count based on workload:

```{r}
# Start with a few daemons
mirai_init(n = 4)

# Scale up for intensive processing
mirai_scale(16)
results <- collect(heavy_flow)

# Scale down when done
mirai_scale(4)

# Check status
mirai_status()
mirai_dispatcher_status()
```

### Chunking Strategies

Optimize task distribution with mirai's dispatcher:

```{r}
# Fine-grained parallelism (many small tasks)
dist_mirai(
  n = 16,
  dispatcher = TRUE,     # Essential for load balancing
  chunks_per_job = 1     # One group per task
)

# Coarse-grained parallelism (fewer large tasks)
dist_mirai(
  n = 8,
  dispatcher = FALSE,    # Direct task assignment
  chunks_per_job = 10    # Bundle groups together
)
```

## Practical Examples

### Example 1: Large-scale simulation on local machine

```{r}
library(parade)

# Parameter grid with 10,000 combinations
grid <- expand.grid(
  alpha = seq(0.1, 1, by = 0.1),
  beta = seq(0.5, 5, by = 0.5),
  n_sim = 1:100
)

# Use mirai to bypass connection limits
fl <- flow(grid) |>
  stage("simulate", function(alpha, beta, n_sim) {
    set.seed(n_sim)
    data <- rgamma(1000, shape = alpha, rate = beta)
    list(
      mean = mean(data),
      var = var(data),
      ks_p = ks.test(data, "pgamma", alpha, beta)$p.value
    )
  }, schema = returns(mean = dbl(), var = dbl(), ks_p = dbl())) |>
  distribute(dist_mirai(n = 200))  # 200 workers!

results <- collect(fl)
```

### Example 2: Remote execution on HPC cluster

```{r}
# Development: test locally
fl_dev <- fl |> 
  distribute(use_mirai_local(n = 4))

test_results <- fl_dev |> collect(limit = 100)

# Production: scale to cluster
fl_prod <- fl |>
  distribute(use_mirai_slurm(
    n = 64,
    partition = "compute",
    time = "12:00:00",
    mem = "256G",
    account = "project-123"
  ))

handle <- submit(fl_prod)

# Monitor progress
while (!all(deferred_status(handle)$done)) {
  status <- deferred_status(handle)
  print(status)
  Sys.sleep(30)
}

final_results <- deferred_collect(handle)
```

### Example 3: Cross-platform workflow

```{r}
# Detect environment and choose appropriate backend
get_distribution <- function() {
  if (Sys.getenv("SLURM_JOB_ID") != "") {
    # Running on SLURM
    use_mirai_slurm(n = 32, partition = "compute")
  } else if (file.exists("~/.ssh/compute_nodes")) {
    # Has SSH access to compute nodes
    nodes <- readLines("~/.ssh/compute_nodes")
    use_mirai_ssh(remotes = nodes, tunnel = TRUE)
  } else {
    # Local development
    use_mirai_local()
  }
}

# Portable workflow
fl |> 
  distribute(get_distribution()) |>
  collect()
```

## Performance Considerations

### When to use mirai vs traditional backends

| Use Case | Best Choice | Reasoning |
|----------|------------|-----------|
| Few large tasks | `dist_local()` or `dist_slurm()` | Traditional backends work well |
| Many small tasks | `dist_mirai()` | Lower per-task overhead |
| >125 parallel tasks | `dist_mirai()` | Only option beyond connection limit |
| Need load balancing | `dist_mirai(dispatcher = TRUE)` | Automatic work distribution |
| Firewalled cluster | `use_mirai_ssh(tunnel = TRUE)` | SSH tunneling support |
| Require TLS security | `dist_mirai(tls = TRUE)` | Built-in encryption |

### Optimizing dispatcher performance

The dispatcher provides automatic load balancing but adds slight overhead:

```{r}
# For homogeneous tasks (similar runtime)
dist_mirai(n = 16, dispatcher = FALSE)

# For heterogeneous tasks (varying runtime)
dist_mirai(n = 16, dispatcher = TRUE)

# For embarrassingly parallel workflows
dist_mirai(
  n = 32,
  dispatcher = TRUE,
  chunks_per_job = 1  # Maximum parallelism
)
```

## Troubleshooting

### Common issues and solutions

**Issue**: "Cannot connect to daemons"
```{r}
# Check daemon status
mirai_status()

# Restart daemons
mirai_stop()
mirai_init()
```

**Issue**: "Tasks not distributing evenly"
```{r}
# Ensure dispatcher is enabled
dist_mirai(n = 16, dispatcher = TRUE)

# Check dispatcher status
mirai_dispatcher_status()
```

**Issue**: "SSH connection fails"
```{r}
# Test SSH manually first
system("ssh node1 echo 'connected'")

# Use verbose mode for debugging
Sys.setenv(MIRAI_DEBUG = "TRUE")
```

**Issue**: "SLURM daemons not starting"
```{r}
# Check SLURM submission manually
mirai::daemons(
  n = 1,
  url = mirai::host_url(tls = TRUE),
  remote = mirai::cluster_config(
    command = "sbatch",
    options = "#SBATCH --partition=debug\n#SBATCH --time=00:05:00"
  )
)

# Check SLURM queue
system("squeue -u $USER")
```

## Best Practices

1. **Always cleanup daemons**: Use `stop_on_exit = TRUE` (default) or explicitly call `mirai_stop()`

2. **Choose appropriate daemon count**: 
   - Local: number of cores
   - Remote: 2-4x number of physical machines
   - SLURM: Based on partition limits

3. **Use TLS for production**: Enable encryption for sensitive data
   ```{r}
   dist_mirai(n = 16, tls = TRUE, port = 5555)
   ```

4. **Monitor resource usage**: Track daemon performance
   ```{r}
   # During execution
   mirai_dispatcher_status()
   
   # After completion
   mirai_status()
   ```

5. **Test locally before scaling**: Verify workflow correctness
   ```{r}
   # Development
   fl |> distribute(use_mirai_local(n = 2)) |> collect(limit = 10)
   
   # Production
   fl |> distribute(use_mirai_slurm(n = 64)) |> submit()
   ```

## Summary

The mirai backend extends parade's capabilities with:

- **Unlimited parallelism**: No connection limits
- **Flexible deployment**: Local, SSH, or SLURM
- **Enhanced security**: TLS and SSH tunneling
- **Better performance**: Low-latency task execution
- **Simple interface**: Drop-in replacement for existing backends

Whether you're developing locally or scaling to hundreds of nodes, mirai provides the performance and flexibility needed for modern parallel computing workflows.
